---
title: "Correlation_heatmap"
output: html_document
date: "2024-09-02"
---

Heatmap: https://r-graph-gallery.com/heatmap.html


```{r}

# Define APA theme
theme_apa <- function(base_size = 12, base_family = "serif") {
  theme_classic(base_size = base_size, base_family = base_family) +
    theme(
      # Title and subtitle
      plot.title = element_text(face = "bold", size = base_size * 1.2, hjust = 0.5),
      plot.subtitle = element_text(size = base_size, hjust = 0.5),
      
      # Axis titles and text
      axis.title = element_text(size = base_size * 1.1),
      axis.text = element_text(size = base_size),
      axis.line = element_line(color = "black"),
      axis.ticks = element_line(color = "black"),
      
      # Panel and background
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      
      # Legend
      legend.title = element_text(size = base_size),
      legend.text = element_text(size = base_size),
      legend.position = "bottom",
      legend.key = element_blank()
    )
}
```

## Dendrogram and Reordering
```{r}
library(ggplot2)
# The mtcars dataset:
data <- as.matrix(mtcars)

# Default Heatmap
heatmap(data)

# Use 'scale' to normalize
heatmap(data, scale="column")

# No dendrogram nor reordering for neither column or row
heatmap(data, Colv = NA, Rowv = NA, scale="column") +  
  
  theme_apa()
```

## Color palette

```{r}
# 1: native palette from R
heatmap(data, scale="column", col = cm.colors(256))  + theme_apa()
heatmap(data, scale="column", col = terrain.colors(256))  + theme_apa()
 
# 2: Rcolorbrewer palette
library(RColorBrewer)
coul <- colorRampPalette(brewer.pal(8, "PiYG"))(25)
heatmap(data, scale="column", col = coul) + theme_apa()
```


# Custom Layout

```{r}
# Add classic arguments like main title and axis title
heatmap(data, Colv = NA, Rowv = NA, scale="column", col = coul, xlab="variable", ylab="car", main="heatmap")
 
# Custom x and y labels with cexRow and labRow (col respectively)
heatmap(data, scale="column", cexRow=1.5, labRow=paste("new_", rownames(data),sep=""), col= colorRampPalette(brewer.pal(8, "Blues"))(25))
```

## Add color beside heatmap

```{r}
# Example: grouping from the first letter:
my_group <- as.numeric(as.factor(substr(rownames(data), 1 , 1)))
colSide <- brewer.pal(9, "Set1")[my_group]
colMain <- colorRampPalette(brewer.pal(8, "Blues"))(25)
heatmap(data, Colv = NA, Rowv = NA, scale="column" , RowSideColors=colSide, col=colMain   )
```

## Using geom_tile() from ggplot2

```{r}
# Library
library(ggplot2)

# Dummy data
x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5)
 
# Heatmap 
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile()
```

# use color palettes: 

```{r}
# Library
library(ggplot2)
library(hrbrthemes)

# Dummy data
x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5)
 
# Give extreme colors:
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile() +
  scale_fill_gradient(low="white", high="blue") +
  theme_ipsum()

# Color Brewer palette
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile() +
  scale_fill_distiller(palette = "RdPu") +
  theme_ipsum()

# Color Brewer palette
library(viridis)
ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile() +
  scale_fill_viridis(discrete=FALSE) +
  theme_ipsum()
```
## From wide input format

```{r}
# Library
library(ggplot2)
library(tidyr)
library(tibble)
library(hrbrthemes)
library(dplyr)

# Volcano dataset
#volcano

# Heatmap 
volcano %>%
  
  # Data wrangling
  as_tibble() %>%
  rowid_to_column(var="X") %>%
  gather(key="Y", value="Z", -1) %>%
  
  # Change Y to numeric
  mutate(Y=as.numeric(gsub("V","",Y))) %>%

  # Viz
  ggplot(aes(X, Y, fill= Z)) + 
    geom_tile() +
    theme_ipsum() +
    theme(legend.position="none")
```

## Turn it interactive with plotly

```{r}
# Library
library(ggplot2)
library(hrbrthemes)
library(plotly)

# Dummy data
x <- LETTERS[1:20]
y <- paste0("var", seq(1,20))
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5)
 
# new column: text for tooltip:
data <- data %>%
  mutate(text = paste0("x: ", x, "\n", "y: ", y, "\n", "Value: ",round(Z,2), "\n", "What else?"))

# classic ggplot, with text in aes
p <- ggplot(data, aes(X, Y, fill= Z, text=text)) + 
  geom_tile() +
  theme_ipsum()

ggplotly(p, tooltip="text")

# save the widget
# library(htmlwidgets)
# saveWidget(pp, file=paste0( getwd(), "/HtmlWidget/ggplotlyHeatmap.html"))
```


## Interactive heatmaps from R: https://www.data-to-viz.com/graph/heatmap.html

## Interactive heatmaps from time series 



```{r}
# Load required libraries
library(ggplot2)
library(dplyr) # easier data wrangling 
library(viridis) # colour blind friendly palette, works in B&W also
library(lubridate) # for easy date manipulation
library(ggExtra) # for additional ggplot theme options
library(tidyr) 

# Simulate a similar dataset (since Interpol.T is no longer available)
set.seed(123)
# Create a sample dataset
df <- data.frame(
  stationid = rep("T0001", 24 * 365),
  date = rep(seq.Date(from = as.Date("2023-01-01"), to = as.Date("2023-12-31"), by = "day"), each = 24),
  hour = rep(0:23, 365),
  temp = rnorm(24 * 365, mean = 10, sd = 5)
)

# Clean up the data
df <- df %>%
  mutate(
    year = year(date),
    month = month(date, label = TRUE),
    day = day(date)
  )

# Fill missing temperature values (for demonstration)
df <- df %>%
  fill(temp)

# Plotting starts here
p <- ggplot(df, aes(day, hour, fill = temp)) +
  geom_tile(color = "white", size = 0.1) + 
  scale_fill_viridis(name = "Hrly Temps C", option = "C") +
  facet_grid(year ~ month) +
  scale_y_continuous(trans = "reverse", breaks = unique(df$hour)) +
  scale_x_continuous(breaks = c(1, 10, 20, 31)) +
  theme_minimal(base_size = 8) +
  labs(title = paste("Hourly Temps - Station", unique(df$stationid)), x = "Day", y = "Hour Commencing") +
  theme(
    legend.position = "bottom",
    plot.title = element_text(size = 14, hjust = 0),
    axis.text.y = element_text(size = 6),
    strip.background = element_rect(colour = "white"),
    axis.ticks = element_blank(),
    axis.text = element_text(size = 7),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 6)
  ) +
  removeGrid()  # from ggExtra

# Display the plot
p

```
Note: https://r-graph-gallery.com/283-the-hourly-heatmap.html packages interpol.T is removed from cran, so the code was adjusted 

## Using the levelplot() function of lattice
```{r}
# Load the lattice package
library("lattice")
 
# Dummy data
x <- seq(1,10, length.out=20)
y <- seq(1,10, length.out=20)
data <- expand.grid(X=x, Y=y)
data$Z <- runif(400, 0, 5)

## Try it out
levelplot(Z ~ X*Y, data=data  ,xlab="X",
          main="")
```


## From wide input matrix

```{r}
# Load the library
library("lattice")
 
# Dummy data
data <- matrix(runif(100, 0, 5) , 10 , 10)
colnames(data) <- letters[c(1:10)]
rownames(data) <- paste( rep("row",10) , c(1:10) , sep=" ")
 
# plot it flipping the axis
levelplot(data)
```

#Flip and reorder axis

```{r}
# Load the library
library("lattice")
 
# Dummy data
data <- matrix(runif(100, 0, 5) , 10 , 10)
colnames(data) <- letters[c(1:10)]
rownames(data) <- paste( rep("row",10) , c(1:10) , sep=" ")
 
# plot it flipping the axis
levelplot( t(data[c(nrow(data):1) , ]),
           col.regions=heat.colors(100))
```

## Custom colors

```{r}
# Lattice package
require(lattice)

# The volcano dataset is provided, it looks like that:
#head(volcano)

# 1: native palette from R
levelplot(volcano, col.regions = terrain.colors(100)) # try cm.colors() or terrain.colors()

# 2: Rcolorbrewer palette
library(RColorBrewer)
coul <- colorRampPalette(brewer.pal(8, "PiYG"))(25)
levelplot(volcano, col.regions = coul) # try cm.colors() or terrain.colors()

# 3: Viridis
library(viridisLite)
coul <- viridis(100)
levelplot(volcano, col.regions = coul) 
#levelplot(volcano, col.regions = magma(100)) 
```
#Heatmap with smoothing - LatticeExtra

```{r}
# library
library(latticeExtra) 
 
# create data
set.seed(1) 
data <- data.frame(x = rnorm(100), y = rnorm(100)) 
data$z <- with(data, x * y + rnorm(100, sd = 1)) 
 
# showing data points on the same color scale 
levelplot(z ~ x * y, data, 
          panel = panel.levelplot.points, cex = 1.2
    ) + 
    layer_(panel.2dsmoother(..., n = 200))


```

