---
title: "Correlation_scatterplot"
output: html_document
date: "2024-08-31"
---

Scatter Plot: https://r-graph-gallery.com/scatterplot.html

```{r}

# Define APA theme
theme_apa <- function(base_size = 12, base_family = "serif") {
  theme_classic(base_size = base_size, base_family = base_family) +
    theme(
      # Title and subtitle
      plot.title = element_text(face = "bold", size = base_size * 1.2, hjust = 0.5),
      plot.subtitle = element_text(size = base_size, hjust = 0.5),
      
      # Axis titles and text
      axis.title = element_text(size = base_size * 1.1),
      axis.text = element_text(size = base_size),
      axis.line = element_line(color = "black"),
      axis.ticks = element_line(color = "black"),
      
      # Panel and background
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      panel.background = element_blank(),
      
      # Legend
      legend.title = element_text(size = base_size),
      legend.text = element_text(size = base_size),
      legend.position = "bottom",
      legend.key = element_blank()
    )
}
```


##Basic scatterplot with R and ggplot2

```{r}


# library
library(ggplot2)
 
# The iris dataset is provided natively by R
#head(iris)
 
# basic scatterplot
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point() +theme_apa()


```



##Custom scatterplot

```{r}


# library
library(ggplot2)
 
# Iris dataset is natively provided by R
#head(iris)
 
# use options!
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point(
        color="orange",
        fill="#69b3a2",
        shape=21,
        alpha=0.5,
        size=6,
        stroke = 2
        ) +theme_apa()

```
##Using theme_ipsum

```{r}


# library
library(ggplot2)
library(hrbrthemes)

# Iris dataset is natively provided by R
#head(iris)
 
# use options!
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width)) + 
    geom_point(
        color="black",
        fill="#69b3a2",
        shape=22,
        alpha=0.5,
        size=6,
        stroke = 1
        ) +
    theme_ipsum() +theme_apa()

```

##aesthetics

```{r}
library(ggplot2)
# Transparency
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, alpha=Species)) + 
    geom_point(size=6, color="#69b3a2") +
    ggtitle("Transparency") +
    theme_apa()

# Shape
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, shape=Species)) + 
    geom_point(size=6, color="lightblue") +
    ggtitle("Shape") +
    theme_apa()

# Size
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, size=Petal.Width)) + 
    geom_point(color="darkred") +
    ggtitle("Size") +
     theme_apa()
```


##Mapping to several features

```{r}
# load ggplot2
library(ggplot2)
library(hrbrthemes)

# A basic scatterplot with color depending on Species
ggplot(iris, aes(x=Sepal.Length, y=Sepal.Width, 
                 shape=Species, 
                 alpha=Species, 
                 size=Species, 
                 color=Species)) + 
    geom_point() +
    theme_ipsum()
```
##  text labels with ggplot2

```{r}


# library
library(ggplot2)
 
# Keep 30 first rows in the mtcars natively available dataset
data=head(mtcars, 30)
 
# 1/ add text with geom_text, use nudge to nudge the text
ggplot(data, aes(x=wt, y=mpg)) +
  geom_point() + # Show dots
  geom_text(
    label=rownames(data), 
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )


```

##Add labels with geom_label()

```{r}


# library
library(ggplot2)
 
# Keep 30 first rows in the mtcars natively available dataset
data=head(mtcars, 30)
 
# 1/ add text with geom_text, use nudge to nudge the text
ggplot(data, aes(x=wt, y=mpg)) +
  geom_point() + # Show dots
  geom_label(
    label=rownames(data), 
    nudge_x = 0.25, nudge_y = 0.25, 
    check_overlap = T
  )


```

##Add one text label only


```{r}
# Keep 30 first rows in the mtcars natively available dataset
data=head(mtcars, 30)
 
# Add one annotation
ggplot(data, aes(x=wt, y=mpg)) +
  geom_point() + # Show dots
  geom_label(
    label="This point", 
    x=4.1,
    y=20,
    label.padding = unit(0.55, "lines"), # Rectangle size around label
    label.size = 0.35,
    color = "red",
    fill="#69b3a2"
  ) +theme_apa()
```

## Add labels for a selection of marker


```{r}


# library
library(ggplot2)
library(dplyr)
library(tibble)

# Keep 30 first rows in the mtcars natively available dataset
data=head(mtcars, 30)

# Change data rownames as a real column called 'carName'
data <- data %>%
  rownames_to_column(var="carName")
  
# Plot
ggplot(data, aes(x=wt, y=mpg)) +
  geom_point() + 
  geom_label( 
    data=data %>% filter(mpg>20 & wt>3), # Filter data first
    aes(label=carName)
  ) +theme_apa()


```

#Marginal distribution with ggplot2 and ggExtra

```{r}
# library
library(ggplot2)
library(ggExtra)
 
# The mtcars dataset is proposed in R
head(mtcars)
 
# classic plot :
p <- ggplot(mtcars, aes(x=wt, y=mpg, color=cyl, size=cyl)) +
      geom_point() +
      theme(legend.position="none")
 
# with marginal histogram
p1 <- ggMarginal(p, type="histogram")
 
# marginal density
p2 <- ggMarginal(p, type="density")
 
# marginal boxplot
p3 <- ggMarginal(p, type="boxplot")

p
p1
p2
p3
```

```{r}


# library
library(ggplot2)
library(ggExtra)
 
# The mtcars dataset is proposed in R
head(mtcars)
 
# classic plot :
p <- ggplot(mtcars, aes(x=wt, y=mpg, color=cyl, size=cyl)) +
      geom_point() +
      theme(legend.position="none")
 
# Set relative size of marginal plots (main plot 10x bigger than marginals)
p1 <- ggMarginal(p, type="histogram", size=10)
 
# Custom marginal plots:
p2 <- ggMarginal(p, type="histogram", fill = "slateblue", xparams = list(  bins=10))
 
# Show only marginal plot for x axis
p3 <- ggMarginal(p, margins = 'x', color="purple", size=4)

p
p1
p2
p3

```

##Linear model and confidence interval

```{r}


# Library
library(ggplot2)
library(hrbrthemes)

# Create dummy data
data <- data.frame(
  cond = rep(c("condition_1", "condition_2"), each=10), 
  my_x = 1:100 + rnorm(100,sd=9), 
  my_y = 1:100 + rnorm(100,sd=16) 
)

# Basic scatter plot.
p1 <- ggplot(data, aes(x=my_x, y=my_y)) + 
  geom_point( color="#69b3a2") +
  theme_ipsum()
 
# with linear trend
p2 <- ggplot(data, aes(x=my_x, y=my_y)) +
  geom_point() +
  geom_smooth(method=lm , color="red", se=FALSE) +
  theme_ipsum()

# linear trend + confidence interval
p3 <- ggplot(data, aes(x=my_x, y=my_y)) +
  geom_point() +
  geom_smooth(method=lm , color="red", fill="#69b3a2", se=TRUE) +
  theme_ipsum()

p
p1
p2
p3
```
##Scatter plot with trend lines and labels on top

```{r}
# Load the package
library(ggplot2)
library(geomtextpath)
```


```{r}
library(hrbrthemes)
data(iris)

ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) +
  geom_point() +
  geom_labelsmooth(aes(label = 'This is the trend line'), fill = "white",
                method = "lm", formula = y ~ x,
                size = 6, linewidth = 2, boxlinewidth = 0.6) +
  theme_bw() + guides(color = 'none') # remove legend
```
##Multiple groups trend lines with labels

```{r}
library(hrbrthemes)
data(iris)

ggplot(iris, aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point() +
  geom_labelsmooth(aes(label = Species), fill = "white",
                method = "lm", formula = y ~ x,
                size = 3, linewidth = 1, boxlinewidth = 0.4) +
  theme_bw() + guides(color = 'none') # remove legend
```


### Scatterplot with based R: https://r-graph-gallery.com/scatterplot.html